\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-09 paper.tex
model removwed

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD to \kD, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$y_\text{max}$ & maximum value of the scaled second embedding component\\
$r_1$, $r_2$ & range of the first and  second embedding component\\
$h_b$, $w_b$ & height and width of the hexagon\\
\addlinespace
$a_r$ & aspect ratio of the 2D layout before scaling\\
$h_r$ & hexagon ratio\\
$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
$q_x$, $q_y$ & buffer amount along the x and y axes\\
\addlinespace
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
\addlinespace
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

We define the NLDR model as function
\(g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k\), with
(hyper-)parameters \(\mathbfit{\theta}\). The parameters,
\(\mathbfit{\theta}\), depend on the choice of \(g\), and can be
considered part of model fitting in the traditional sense. Common
choices for \(g\) include functions used in tSNE, UMAP, PHATE, TriMAP,
PaCMAP, or MDS, although in theory any function that maps the entire
space \(\mathbb{R}^p\) to a subset of \(\mathbb{R}^k\) is suitable.

>>>>

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
$\mathbfit{y}$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$g$ & functional mapping from \pD to \kD, especially as prescribed by NLDR\\
$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
$y_\text{max}$ & maximum value of the scaled second embedding component\\
$r_1$, $r_2$ & range of the first and  second embedding component\\
$h_b$, $w_b$ & height and width of the hexagon\\
\addlinespace
$a_r$ & aspect ratio of the 2D layout before scaling\\
$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
$q_x$, $q_y$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
\addlinespace
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
\addlinespace
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

We define the NLDR as a function
\(g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k\), with
(hyper-)parameters \(\mathbfit{\theta}\). The parameters,
\(\mathbfit{\theta}\), depend on the choice of \(g\), and can be
considered part of model fitting in the traditional sense. Common
choices for \(g\) include functions used in tSNE, UMAP, PHATE, TriMAP,
PaCMAP, or MDS, although in theory any function that maps the entire
space \(\mathbb{R}^p\) to a subset of \(\mathbb{R}^k\) is suitable.
